---
title: "ComparingHIndex"
output: html_document
---

The purpose of this document is to show that computing the sum of the normalixed effects for each gene-snp pair is more accurate using MASH then ASH. First, recall the RMSE result:


```{r load.data}
data=readRDS("~/Dropbox/simdata.rds")
t=data$tstat;bhat=data$betahat;sebetahat=data$sebetahat;beta=data$beta;v.j=matrix(rep(1,ncol(t)*nrow(t)),ncol=ncol(t),nrow=nrow(t))
covmat=readRDS("~/Dropbox/covmattestim.rds")

t.truth.test=beta/sebetahat
mash.means=read.table("~/Dropbox/Simulationsposterior.means.txt")[,-1]
univariate.ash.pm=read.table("~/Dropbox/univariate.ash.pm.txt")
bma.pm=read.table("~/Dropbox/testsimBMAonlyposterior.means.txt")[,-1]

sqrt(mean((t.truth.test[1:10000,]-mash.means)^2))
sqrt(mean((t.truth.test[1:10000,]-univariate.ash.pm)^2))
sqrt(mean((t.truth.test[1:10000,]-bma.pm)^2))
```

In the first simulation of 'het indexes', for each gene-snp pair, I sum the non-negative normalized effects for the true betas, mash betas, and ash betas and compute the RMSE (i.e., $\sqrt(\sum(truth-estimated)^2))$

```{r}
het.index=function(data,thresh){
  t(apply(data,1,function(x){
   maxx=x[which.max(abs(x))]
   if(maxx==0){h=44}
   else{
     normx=x/x[which.max(abs(x))]
     pos=normx[normx>0]
     h=sum(pos)}
   return(h)
     
    }))}



true.h=het.index(beta[1:10000,],thresh=0.5)
mash.h=het.index(mash.means*sebetahat[1:10000,],0.5)
ash.h=het.index(univariate.ash.pm*sebetahat[1:10000,],0.5)
bma.h=het.index(bma.pm*sebetahat[1:10000,],0.5)
sqrt(mean(true.h-mash.h)^2)
sqrt(mean(true.h-ash.h)^2)
sqrt(mean(true.h-bma.h)^2)
```

In the second, simulation of 'het indexes', for each gene-snp pair, I count number of normalized effects greater than 50% of the max effect for the pair for the true betas, mash betas, and ash betas and compute the RMSE (i.e., $\sqrt(\sum(truth-estimated)^2))$

```{r het.index.two}
het.index.two=function(data,thresh){
  t(apply(data,1,function(x){
   maxx=x[which.max(abs(x))]
   if(maxx==0){h=44}
   else{
     normx=x/x[which.max(abs(x))]
     
     h=sum(normx>thresh)}
   return(h)
     
    }))}


true.h=het.index.two(beta[1:10000,],thresh=0.5)
mash.h=het.index.two(mash.means*sebetahat[1:10000,],0.5)
ash.h=het.index.two(univariate.ash.pm*sebetahat[1:10000,],0.5)
bma.h=het.index.two(bma.pm*sebetahat[1:10000,],0.5)

sqrt(mean(true.h-mash.h)^2)
sqrt(mean(true.h-ash.h)^2)
sqrt(mean(true.h-bma.h)^2)
```


I can also do this for the tissue specific examples. For each gene-snp pair, I count number of normalized effects greater than 50% of the max effect for the pair for the true betas, mash betas, and ash betas and compute the RMSE (i.e., $\sqrt(\sum(truth-estimated)^2))$

```{r het.index.two.tspec}
rm(data)
data=readRDS("~/Dropbox/simdata_tspec.rds")
t=data$tstat;bhat=data$betahat;sebetahat=data$sebetahat;beta=data$beta;v.j=matrix(rep(1,ncol(t)*nrow(t)),ncol=ncol(t),nrow=nrow(t))


t.truth.test=beta/sebetahat
mash.means=read.table("~/Dropbox/testtissuespecposterior.means.txt")[,-1]
univariate.ash.pm=read.table("~/Dropbox/univariate.ash.pm.tspec.txt")
bma.pm=read.table("~/Dropbox/testimBMAonly_tspecposterior.means.txt")[,-1]

true.h=het.index.two(beta[1:10000,],thresh=0.5)
mash.h=het.index.two(mash.means*sebetahat[1:10000,],0.5)
ash.h=het.index.two(univariate.ash.pm*sebetahat[1:10000,],0.5)
bma.h=het.index.two(bma.pm*sebetahat[1:10000,],0.5)

sqrt(mean(true.h-mash.h)^2)
sqrt(mean(true.h-ash.h)^2)
sqrt(mean(true.h-bma.h)^2)
```



